\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\title{tcl-autonomous-mechanism-paper}
\author{Christopher Gardner \\ TAR (Thermodynamic Autonomous Researcher)}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This manuscript has been opened automatically by TAR's planning loop.
Current readiness: outline\_now. Frontier linkage: fp-regime-detection-accuracy.
Director recommendation: Anchor every manuscript in the external problem 'Reliable Long-Horizon Learning Without Catastrophic Forgetting'. Present TAR/TCL/ASC as TAR's unpublished internal work under evaluation, never as established literature or an already-proven solution. Reject unrelated side topics, and frame every section around the concrete problem TAR chose to solve.
\end{abstract}

\section{Planning Status}
This paper has been started from a Research Director directive so the project registry,
author queue, and dashboard all point at the same manuscript target.

\paragraph{Waiting on experiments.}
director-catastrophic-forgetting-boundary-reset-probe, director-catastrophic-forgetting-carryover-probe, regime\_detection\_harder\_dataset\_cifar100, regime\_detection\_mlp\_backbone\_permuted\_mnist

\paragraph{Next action.}
Anchor every manuscript in the external problem 'Reliable Long-Horizon Learning Without Catastrophic Forgetting'. Present TAR/TCL/ASC as TAR's unpublished internal work under evaluation, never as established literature or an already-proven solution. Reject unrelated side topics, and frame every section around the concrete problem TAR chose to solve.

\end{document}
