\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\title{main\_tcl\_scaleup\_paper}
\author{Christopher Gardner \\ TAR (Thermodynamic Autonomous Researcher)}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This manuscript has been opened automatically by TAR's planning loop.
Current readiness: outline\_now. Frontier linkage: fp-scale-up.
Director recommendation: Anchor every manuscript in the external problem 'Scalable Continual Adaptation on Realistic Visual Streams'. Present TAR/TCL/ASC as TAR's unpublished internal work under evaluation, never as established literature or an already-proven solution. Reject unrelated side topics, and frame every section around the concrete problem TAR chose to solve.
\end{abstract}

\section{Planning Status}
This paper has been started from a Research Director directive so the project registry,
author queue, and dashboard all point at the same manuscript target.

\paragraph{Waiting on experiments.}
backbone\_scaling\_vit\_tiny\_cifar100, stream\_difficulty\_progression\_tinyimagenet\_resnet34

\paragraph{Next action.}
Anchor every manuscript in the external problem 'Scalable Continual Adaptation on Realistic Visual Streams'. Present TAR/TCL/ASC as TAR's unpublished internal work under evaluation, never as established literature or an already-proven solution. Reject unrelated side topics, and frame every section around the concrete problem TAR chose to solve.

\end{document}
