% s1_introduction.tex
% Status: DRAFT SCAFFOLD — no claims finalised, no citations verified
% Last updated: 2026-04-19 (scaffold only)
% Lock condition: after Phase 10 aggregate stats confirmed + SI robustness appendix drafted

\section{Introduction}
\label{sec:intro}

% HOOK — methodological observation (Contribution 3)
% PLACEHOLDER — tighten after results locked

Continual learning methods are evaluated primarily on \emph{forgetting}: the degree to
which learning a new task degrades performance on previous tasks.
We show that this metric has a systematic failure mode.
Importance-weighted regularisation methods --- specifically Synaptic Intelligence (SI)
\needsref{Zenke2017} --- can achieve low forgetting scores while producing networks
that have not learned to distinguish classes at all.
A network predicting the same class for every input does not forget what it never learned.

% CLAIM SUMMARY — matches evidence in hand
We introduce the Thermodynamic Continual Learning (TCL) governor, a method that uses
per-layer activation entropy as a signal to regulate weight consolidation during
sequential task learning.
Across five random seeds on Split-CIFAR-10 with a ResNet-18 backbone, TCL achieves
lower forgetting than EWC \needsref{Kirkpatrick2017} with comparable or higher accuracy,
and is the only method in our comparison that simultaneously achieves low forgetting
\emph{and} meaningful learned representations.

% THREE CONTRIBUTIONS — locked, do not expand without discussion
\noindent\textbf{Contributions:}
\begin{enumerate}
    \item \textbf{TCL method.} A thermodynamic governor for continual learning based on
          per-task activation entropy anchoring and dimensionality-weighted L2 regularisation.
          A pre-registered ablation shows neither component reliably reduces
          forgetting in isolation; the combination significantly outperforms SGD
          and each component individually, with substantially reduced variance
          across seeds.
          A complementary $\lambda$ sweep confirms that TCL matches best-case
          EWC mean forgetting without the collapse risk observed at aggressive
          $\lambda$ values (Appendix~\ref{app:ewc_sweep}).

    \item \textbf{Pareto-optimal accuracy-forgetting tradeoff.} TCL dominates EWC on the
          joint accuracy-forgetting (JAF) metric across all five seeds. The gap widens on
          seeds where EWC shows partial accuracy collapse, suggesting TCL degrades more
          gracefully under difficult initialisations.

    \item \textbf{Methodological observation.} Standard forgetting-only rankings can
          misrepresent method quality when importance-weighted methods collapse to
          degenerate solutions. We demonstrate this cleanly on a standard benchmark and
          propose JAF as a minimal two-metric reporting standard.
\end{enumerate}

% PAPER STRUCTURE
The remainder of the paper is organised as follows.
Section~\ref{sec:background} covers relevant prior work.
Section~\ref{sec:method} describes the TCL governor.
Section~\ref{sec:experiments} presents experimental setup and results.
Section~\ref{sec:discussion} discusses limitations and future work.
