% s4_experiments.tex
% Status: COMPLETE — all 5 seeds in, aggregate stats computed, section locked
% Last updated: 2026-04-19 (Phase 10 complete; all seeds 42, 0, 1, 2, 3)
% Lock condition: MET

\section{Experiments}
\label{sec:experiments}

\subsection{Setup}
\label{sec:setup}

\textbf{Dataset.} Split-CIFAR-10: 5 sequential binary classification tasks, 1000
samples per task, no replay buffer. Task identity provided at test time
(task-incremental setting).

\textbf{Backbone.} ResNet-18 ($\sim$11M parameters). Warmup guard: anchor
collection delayed 60 batches to avoid anchoring on random-initialisation noise.
Eigenvalue decomposition disabled (\texttt{compute\_dpr=False}) for efficiency;
regime detection via $\sigma/\sigma^\star$ unchanged.

\textbf{Baselines.}
\begin{itemize}
    \item \textbf{SGD} — fine-tuning with no anti-forgetting mechanism.
    \item \textbf{EWC} \needsref{Kirkpatrick2017} — elastic weight consolidation,
          $\lambda=100$ (selected by sweep in Phase 7; see Appendix~\ref{app:ewc_sweep}).
    \item \textbf{SI} \needsref{Zenke2017} — synaptic intelligence at published default
          hyperparameters ($c=0.1$, $\xi=0.001$). Post-hoc robustness sweep reported in
          Appendix~\ref{app:si_sweep} (labelled post-hoc, not pre-registered).
    \item \textbf{TCL} — thermodynamic governor with dimensionality-weighted L2 penalty,
          $\lambda_{\text{TCL}}=0.01$, $\alpha=0.5$, anchor window 20 batches.
\end{itemize}

\textbf{Protocol.} 5 seeds $\{42, 0, 1, 2, 3\}$, 40 epochs per task, batch size 32.
Pre-registered outcome thresholds: Outcome A $= $ TCL vs EWC mean forgetting delta
$< -0.01$, $p < 0.05$, Cohen's $d > 0.5$.

\textbf{Metrics.} We report:
(1) \emph{Mean forgetting} $\mathcal{F}$: average drop in task accuracy after subsequent
tasks, following \needsref{Lopez-Paz2017}.
(2) \emph{Final mean accuracy} $\mathcal{A}$: mean accuracy across all tasks at end of
training.
(3) \emph{Joint accuracy-forgetting} (JAF): $\mathcal{A} \times (1 - \mathcal{F})$,
a scalar that penalises low-forgetting solutions achieved at the cost of collapsed learning.


\subsection{Results}
\label{sec:results}

\begin{table}[h]
\centering
\caption{Per-seed results on Split-CIFAR-10 (ResNet-18, 40 epochs/task).
Forgetting $\mathcal{F}$ (lower is better), accuracy $\mathcal{A}$ (higher is better),
JAF $= \mathcal{A}\times(1-\mathcal{F})$ (higher is better).
SI acc$=0.50$ indicates collapse to chance-level prediction (2-class tasks).}
\label{tab:main_results}
\begin{tabular}{llccc}
\toprule
Seed & Method & $\mathcal{F}\downarrow$ & $\mathcal{A}\uparrow$ & JAF$\uparrow$ \\
\midrule
\multirow{4}{*}{42}
 & \textbf{TCL} & \textbf{0.1269} & \textbf{0.767} & \textbf{0.669} \\
 & EWC          & 0.1640          & 0.760          & 0.635          \\
 & SI           & 0.1430          & 0.500$^\dagger$& 0.428          \\
 & SGD          & 0.3348          & 0.596          & 0.396          \\
\midrule
\multirow{4}{*}{0}
 & \textbf{TCL} & \textbf{0.1294} & \textbf{0.769} & \textbf{0.670} \\
 & EWC          & 0.1530          & 0.772          & 0.655          \\
 & SI           & 0.1141          & 0.500$^\dagger$& 0.443          \\
 & SGD          & 0.2127          & 0.717          & 0.564          \\
\midrule
\multirow{4}{*}{1}
 & \textbf{TCL} & \textbf{0.1697} & \textbf{0.731} & \textbf{0.607} \\
 & EWC          & 0.2520          & 0.666          & 0.498          \\
 & SI           & 0.1262          & 0.500$^\dagger$& 0.437          \\
 & SGD          & 0.1862          & 0.743          & 0.604          \\
\midrule
\multirow{4}{*}{2}
 & \textbf{TCL} & \textbf{0.1007} & \textbf{0.796} & \textbf{0.716} \\
 & EWC          & 0.2366          & 0.681          & 0.520          \\
 & SI           & 0.1415          & 0.500$^\dagger$& 0.430          \\
 & SGD          & 0.2720          & 0.657          & 0.478          \\
\midrule
\multirow{4}{*}{3}
 & \textbf{TCL} & \textbf{0.1108} & \textbf{0.786} & \textbf{0.699} \\
 & EWC          & 0.1598          & 0.762          & 0.640          \\
 & SI           & 0.1315          & 0.500$^\dagger$& 0.434          \\
 & SGD          & 0.2930          & 0.636          & 0.450          \\
\midrule
\multirow{4}{*}{\textit{Mean$\pm$std}}
 & \textbf{TCL} & \textbf{0.1275$\pm$0.026} & \textbf{0.770$\pm$0.025} & \textbf{0.672$\pm$0.042} \\
 & EWC          & 0.1931$\pm$0.047          & 0.728$\pm$0.050          & 0.590$\pm$0.074          \\
 & SI           & 0.1313$\pm$0.012          & 0.500$\pm$0.000           & 0.434$\pm$0.006          \\
 & SGD          & 0.2597$\pm$0.060          & 0.670$\pm$0.060           & 0.499$\pm$0.085          \\
\bottomrule
\end{tabular}
\\[4pt]
\small $^\dagger$ SI accuracy = 0.500 = random chance on 2-class tasks.
Network has not learned class-discriminative features; low forgetting is an artefact.
\end{table}

\subsection{Statistical Analysis}
\label{sec:stats}

Across all five seeds, TCL achieves lower forgetting than EWC~($\lambda=100$) on every
seed (5/5 wins).
Per-seed forgetting deltas (TCL$-$EWC) are
$-0.037$, $-0.024$, $-0.082$, $-0.136$, $-0.049$ for seeds $42, 0, 1, 2, 3$
respectively; mean delta $= -0.0656$.
A one-sample $t$-test on the five deltas against zero (scipy \texttt{ttest\_1samp})
gives $t(4) = -3.26$, $p = 0.0310$, Cohen's $d = 1.46$.

Pre-registered Outcome~A requires mean delta $< -0.01$, $p < 0.05$, and $d > 0.5$.
All three conditions are satisfied.
\textbf{Outcome~A: confirmed.}

On the JAF metric, TCL dominates EWC across all five seeds (5/5 wins;
mean $0.672 \pm 0.042$ vs.\ $0.590 \pm 0.074$).
The JAF gap is largest on seeds where EWC accuracy falls below $0.70$
(seeds~1 and~2; gaps of $0.109$ and $0.196$ respectively),
and narrowest where EWC accuracy remains above $0.70$
(seeds~42, 0, and~3; gaps of $0.034$, $0.016$, and $0.059$).
We interpret this pattern as TCL maintaining more stable accuracy under
initialisations that cause EWC partial collapse, though the underlying mechanism
is not confirmed from these experiments alone.

TCL also reduces forgetting relative to SGD across all five seeds
($t(4) = -3.70$, $p = 0.0208$, $d = 1.66$, 5/5 wins), consistent with the
Phase~9 result on the same benchmark ($p = 0.0113$, $d = 1.99$).

\subsection{The SI Collapse}
\label{sec:si_collapse}

SI reports forgetting scores of $0.114$--$0.143$ across all five seeds, which would
rank it first or second in a forgetting-only comparison.
However, SI accuracy is exactly $0.500$ on every seed ($n=5$) --- the
chance-level baseline for 2-class tasks.
A network predicting one class regardless of input cannot forget: the forgetting metric
measures the \emph{relative} drop in performance, and if task accuracy never exceeds
chance, there is nothing to lose.

This is not a marginal result.
Exactly $0.500$ accuracy (rather than, say, $0.51$ or $0.53$) indicates the network
has converged to a single-class decision boundary.
At SI's published default hyperparameters ($c=0.1$, $\xi=0.001$), the importance
weights become sufficiently large to prevent the network from updating meaningfully
after the first task.
This is the failure mode of over-constrained importance weighting, analogous to but
distinct from the EWC collapse we observed at $\lambda=1000$ in our hyperparameter
sweep (Appendix~\ref{app:ewc_sweep}).

We did not re-tune SI hyperparameters for this comparison because TCL is the method
under evaluation.
A post-hoc robustness sweep (Appendix~\ref{app:si_sweep}) tests SI at $c \in
\{0.01, 0.1, 0.5\}$ to establish whether non-degenerate SI configurations exist on
this benchmark and, if so, whether TCL remains competitive.


\subsection{Ablation Study}
\label{sec:ablation}

To identify which components of TCL drive the forgetting reduction, we ran a
pre-registered four-condition ablation on the same benchmark (ResNet-18, 5 seeds
$\{42, 0, 1, 2, 3\}$, 40 epochs/task):

\begin{itemize}
    \item \textbf{SGD} — fine-tuning baseline (no anti-forgetting mechanism).
    \item \textbf{Governor-only} — thermodynamic LR adjustment active; L2 penalty
          disabled ($\lambda_{\text{TCL}}=0$).
    \item \textbf{Penalty-only} — D$_{\mathrm{PR}}$-weighted L2 anchor active at
          $D_{\mathrm{PR}}=1.0$ (no governor); observer disabled.
    \item \textbf{Full TCL} — both components active (Phase~10 configuration).
\end{itemize}

\begin{table}[h]
\centering
\caption{Ablation results on Split-CIFAR-10 (ResNet-18, 40 epochs/task).
Forgetting $\mathcal{F}$ (lower is better). Per-seed and aggregate (mean$\pm$std).}
\label{tab:ablation}
\begin{tabular}{lcccccc}
\toprule
Condition & Seed 42 & Seed 0 & Seed 1 & Seed 2 & Seed 3 & Mean$\pm$std \\
\midrule
SGD             & 0.2471 & 0.1720 & 0.1495 & 0.2601 & 0.3485 & 0.2354$\pm$0.079 \\
Governor-only   & 0.2107 & 0.2175 & 0.2771 & 0.2047 & 0.2669 & 0.2354$\pm$0.034 \\
Penalty-only    & 0.1485 & 0.1909 & 0.1393 & 0.3016 & 0.1720 & 0.1905$\pm$0.065 \\
\textbf{Full TCL} & \textbf{0.1206} & \textbf{0.1083} & \textbf{0.1180} & \textbf{0.1225} & \textbf{0.1044} & \textbf{0.1148$\pm$0.008} \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Component analysis.}
Governor-only produces identical mean forgetting to SGD
($\Delta = -0.0001$, $p = 0.999$, $d = 0.001$, $3/5$ seeds).
The LR adjustment alone provides no reduction in forgetting.
Penalty-only reduces forgetting by $0.045$ relative to SGD on average
($\Delta = -0.045$, $p = 0.330$, $d = 0.496$, $3/5$ seeds), a directional
improvement that does not reach significance across five seeds and is
accompanied by high variance (std $= 0.065$; seed~2 forgetting $= 0.302$,
worse than all other conditions on that seed).

Full TCL is better than SGD on all five seeds
($\Delta = -0.121$, $t(4) = -3.30$, $p = 0.030$, $d = 1.48$, $5/5$)
and better than governor-only on all five seeds
($\Delta = -0.121$, $t(4) = -7.10$, $p = 0.002$, $d = 3.18$, $5/5$).
The comparison with penalty-only is directional ($\Delta = -0.076$, $d = 1.20$,
$5/5$ seeds) but does not reach significance at $\alpha = 0.05$ ($p = 0.056$).
Notably, full TCL has substantially lower forgetting variance than penalty-only
alone (std $= 0.008$ vs.\ $0.065$), an observation consistent with the governor
stabilising the penalty's effect, though the mechanism is not confirmed from
these data alone.

\textbf{Interpretation.}
The pre-registered directional classifier returned \textsc{Penalty-Dominant}
because penalty-only's mean delta exceeded the $-0.01$ threshold while
governor-only's did not.
The statistics are more constraining: \emph{neither component alone reduces
forgetting significantly} (both $p > 0.05$ vs.\ SGD at $n = 5$).
Only the combined method clears significance ($p = 0.030$).
The full TCL system's advantage over penalty-only approaches but does not
reach significance ($p = 0.056$, $d = 1.20$), and would require larger $n$
to resolve.

The most statistically clear signal is in variance.
Full TCL's seed-to-seed forgetting std is $0.008$, compared to $0.065$
for penalty-only, $0.034$ for governor-only, and $0.079$ for SGD.
Penalty-only can actively worsen forgetting relative to SGD
(seed~2: $\mathcal{F} = 0.302$ vs.\ $0.260$); full TCL never did.
This pattern is consistent with the governor preventing the penalty's
worst-case behaviour and stabilising its effect across random initialisations,
though the precise mechanism is not confirmed from these data alone.

The accurate claim supported by this ablation: the L2 anchor provides the
directional benefit; the regime governor provides the consistency.
Together they produce a method that significantly outperforms SGD with
qualitatively tighter variance than either component in isolation.
Contribution~1 is supported with this qualification.
